Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-23 paper/paper.tex
updated references

% applications of cvnn
Complex-valued nature of the data in acoustic and radio signal processing has been
the main driver behind the adoption of $\cplx$-valued neural networks ($\cplx$VNN).
\citet{hirose_complex-valued_2009} argues that the combined phase-magnitude effect of
$\cplx$-valued transformations removes the excess degrees of freedom, that cause
degenerate transformations in $\real$-valued networks with twice the feature dimensions.
Their study demonstrates superiority of $\cplx$VNN in landmine detection using ground
penetrating radar imaging. Other examples, where $\cplx$-valued networks have outperformed
$\real$-valued networks, include magnetic resonance \citep{hui_mri_1995,wang_deepcomplexmri_2020}
and radar imaging \citep{haensch_complex-valued_2010,zhang_complex-valued_2017}, music
transcription and spectral speech modelling \citep{wisdom_full-capacity_2016,trabelsi_deep_2018},
and wireless signal classification \citep{yang_complex_2019}. \citet{tarver_design_2019}
have lowered the out-of-band power leakage with a $\cplx$-valued network for digital
signal predistortion. The networks have also been applied to non-$\cplx$-valued domains,
such as image classification \citep{popa_complex-valued_2017}, sequence modelling
\citep{danihelka_associative_2016}, and motion prediction \citep{wolter_complex_2018},
and for stabilizing back-propagation in RNN \citep{wisdom_full-capacity_2016}.

@@

\citet{kharitonov_variational_2018} addressed theoretical issues with improper prior
$\pi$ in Sparse VD, emphasized by \citet{hron_variational_2018}, and proposed
\emph{Automatic Relevance Determination} Variational Dropout, by replacing $\pi(\omega_j)$
with a proper Gaussian prior $
  \pi_\lambda(\omega_j) = \mathcal{N}(\omega_j \vert\, 0, \tau_j^{-1})
$ with learnable precision $\tau_j > 0$ \citep{neal_bayesian_1996}. This recast the VD
as the \emph{Empirical Bayes} approach, which performs Bayesian Inference over $\omega$,
but uses Maximum Likelihood estimates for the hyper-parameters $\lambda$,
\citep{mackay_bayesian_1994}. Maximizing \eqref{eq:sgvb_estimator} over $\tau$, holding
other parameters fixed, yields $
  \tau^*_j = {(\mu_j^2 + \sigma^2_j)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  K(\alpha)
    = \frac12 \sum_j \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}
% \begin{align}
%   \pi(\omega)
%     &
%     = \prod_J \pi(s_J)
%       \prod_{j\in J} \pi(\omega_j \vert s_J)
%     \propto \prod_J \frac1{\lvert s_J \rvert}
%       \prod_{j\in J} \mathcal{N}(\omega_j \vert 0, s_J^2)
%     \,, \\
%   q_\theta(\omega)
%     &
%     % = \prod_J q_\theta(s_J)
%     %   \prod_{j\in J} q_\theta(\omega_j \vert s_J)
%     = \prod_J \mathcal{N}(s_J \vert \mu_J, \alpha_J \mu_J^2)
%       \prod_{j\in J} \mathcal{N}(\omega_j \vert \mu_j s_J, \sigma_j^2 s_J^2)
%     \,.
% \end{align}

@@

% Although $\cplx$-valued dense layers and back-propagation have been around since early
% 1990-s, one of the earliest accounts
Development of deep $\cplx$-valued networks has been active. \citet{haensch_complex-valued_2010}
put forward $\cplx$-valued convolutional networks, \citet{guberman_complex_2016} and
\citet{popa_complex-valued_2017} developed modifications of pooling, \citet{arjovsky_unitary_2016}
and \citet{wisdom_full-capacity_2016} proposed $\cplx$-valued RNNs with unitary recurrent
transition matrices, and \citet{danihelka_associative_2016} developed $\cplx$-valued
holographic representations for LSTMs. More recently \citet{trabelsi_deep_2018} proposed
$\cplx$-valued batch-normalization and weight initialization, \citet{wolter_complex_2018}
investigated different $\cplx$-valued gating mechanisms for RNNs, and \citet{yang_complex_2019}
proposed $\cplx$-valued self-attention and complex transformer architecture. It merits
noting that \citet{gaudet_deep_2018} generalized $\cplx$VNN further to deep quaternion-valued
networks, and \citet{vecchi_compressing_2020} studied sparsity inducing regularizers for
them.
% complex-valued distributions \citep{pav_moments_2015,taubock_complex-valued_2012},
% and \citep{karseras_caution:_2014}

@@

% Prior experiments demonstrated that the $\cplx$-valued variational sparsification
% what can be expected from prior research.
MusicNet is a corpus of $330$ annotated classical music recordings used for learning
feature representations for music transcription tasks \citep{thickstun_learning_2017}.
\citet{trabelsi_deep_2018} have proposed a $1d$ VGG-like $\cplx$-valued network that
surpassed a similar $\real$-valued network and achieved $72.9\%$ pooled \emph{Average
Precision} on this dataset. Recently \citet{yang_complex_2019} have reported $74.2\%$
AP with a $\cplx$-valued transformer, \citet{thickstun_invariances_2018} have achieved
$77.3\%$ with a four-layer $\real$-valued network on $\log$-spaced spectrogram, and
\citet{draguns_residual_2020} report $78.0\%$ with a residual-shuffle-exchange network.

@@

In this section we derive the exact gradient of \eqref{eq:improper-kl-div-real}. Even
though, ultimately, the divergence involves a hypergeometric function, and its gradient
requires evaluating the Gauss error function, the analysis provides independent verification
of the approximation, proposed by \citet{molchanov_variational_2017}. The logic of the
derivation follows \citet{lapidoth_capacity_2003}.
% \citet{pav_moments_2015} correctly notice that \cite[p. 2466]{lapidoth_capacity_2003} has a missing $\log{2}$ term.

@@

$, the random variable $W = \sum_i (\mu_i + z_i)^2$ is noncentral $\chi^2$ with shape $m$
and noncentrality parameter $\lambda = \sum_i \mu_i^2$, i.e. $W\sim \chi^2_m(\lambda)$.
Its density is
\begin{equation}  \label{eq:nonchisq-density}
  f_W(x)
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0} \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %   \tfrac{
    %     x^{n + \tfrac{m}2 - 1} e^{-\tfrac{x}2}
    %   }{
    %     2^{n + \tfrac{m}2} \Gamma(n + \tfrac{m}2)
    %   }
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0} \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %     \tfrac{
    %         \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1} e^{-\tfrac{x}2}
    %     }{
    %         2 \Gamma(n + \tfrac{m}2)
    %     }
    = \frac12 e^{- \tfrac{x + \lambda}2} \bigl(\tfrac{x}2\bigr)^{\tfrac{m}2 - 1}
      \sum_{n \geq 0} \frac{
        \bigl(\tfrac{x \lambda}4\bigr)^n
      }{
        n! \Gamma(n + \tfrac{m}2)
      }
    % = \frac12 e^{- \tfrac{x + \lambda}2}
    % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m}4 - \tfrac12}
    % \bigl(\tfrac{x \lambda}4\bigr)^{\tfrac{m}4 - \tfrac12}
    % \sum_{n \geq 0} \tfrac{
    %         \bigl(\tfrac{x \lambda}4\bigr)^n
    %     }{
    %         n! \Gamma(n + \tfrac{m}2)
    %     }
    % = \frac12 e^{- \tfrac{x + \lambda}2}
    % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m}4 - \tfrac12}
    % \bigl(\tfrac{\sqrt{x \lambda}}2\bigr)^{\tfrac{m}2 - 1}
    % \sum_{n \geq 0} \tfrac{
    %         \bigl(\tfrac{x \lambda}4\bigr)^n
    %     }{
    %         n! \Gamma(n + \tfrac{m}2)
    %     }
    % = \frac12 e^{- \tfrac{x + \lambda}2}
    % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m - 2}4}
    % I_{\bigl(\tfrac{m}2 - 1\bigr)}(\sqrt{x \lambda})

@@

% where $I_k$ is the modified Bessel function of the first kind
% I_k(s)
%   = \Bigl(\frac{s}2\Bigr)^k
%       \sum_{n \geq 0} \tfrac{
%           \bigl( \tfrac{s}2 \bigr)^{2n}
%       }{n! \Gamma(n + k + 1)}
By integrating the power series \eqref{eq:nonchisq-density} with substitution $x \to 2u$,
the expectation $
  \mathbb{E}_{W\sim \chi^2_m(\lambda)} \log W
$, needed in \eqref{eq:improper-kl-div-real}, equals
\begin{multline}  \label{eq:exp-log-chisq-proof}
    % &= \int_0^\infty f_W(x) \log x dx
    % &\frac12 e^{- \tfrac\lambda2}
    % \int_0^\infty \sum_{n \geq 0} \biggl(
    %     \frac{
    %       e^{- \tfrac{x}2} \bigl(\tfrac\lambda2\bigr)^n
    %     }{
    %       n! \Gamma(n + \tfrac{m}2)
    %     }
    %     \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1}
    % \biggr) \log{(x)} dx
    % = [\text{ absolute summability and Fubini, or any other conv. thm for integrals of nonnegative integrals}]
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0}
    %     \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %     \tfrac1{\Gamma(n + \tfrac{m}2)}
    % \int_0^\infty
    %     e^{- \tfrac{x}2} \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1}
    %     (\log2 + \log \tfrac{x}2) \tfrac{dx}2
    % = [u = \tfrac{x}2]
    e^{- \tfrac\lambda2} \sum_{n \geq 0}
        \frac{\bigl(\tfrac\lambda2\bigr)^n}{
          n! \Gamma(n + \tfrac{m}2)
        }
    \int_0^\infty
        e^{- u} u^{n + \tfrac{m}2 - 1}
        \log{(2 u)} du
    % = [\text{definitions:} \Gamma, \psi]
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0}
    %     \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    % (\psi(n + \tfrac{m}2) + \log2)
    % \notag
    = \log{2}
    + e^{- \tfrac\lambda2} \sum_{n \geq 0}
        \frac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
        \psi(n + \tfrac{m}2)
\end{multline}
where $\psi(x)$ is the digamma function, i.e.
\begin{equation}  \label{eq:digamma}
  \psi(x)
    % = \frac{d}{dx} \log \Gamma(x)
    = \frac1{\Gamma(x)}
      \int_0^\infty
        u^{x-1} e^{-u} \log{u}
      \, du

@@

which satisfies $
  \psi(z+1) = \psi(z) + \tfrac1z
$, $
  \psi(x)\leq \log x - \tfrac1{2x}
$, and $
  \psi(\tfrac12) = -\gamma - 2\log 2
$ ($\gamma$ is Euler's constant). If we put
\begin{equation}  \label{eq:g-m-series}

@@

    % alpha: e^{-x} \sum_{n=0}^{\infty} \frac{x^n}{n!} \psi(n + m / 2)
    % = e^{-x} \sum_{n \geq 0} \frac{x^n}{n! \Gamma(n + \tfrac{m}2)}
    %     \int_0^\infty e^{-t} t^{n + \tfrac{m}2-1} \log{t} dt
    = e^{-x} \sum_{n \geq 0} \frac{x^n}{n!} \psi(n + \tfrac{m}2)
then the desired expectation in \eqref{eq:improper-kl-div-real} equals $
  \log{2} + g_1(\tfrac1{2\alpha})
% We can differentiate the series in $g_m(x)$, since the sum converges on $\mathbb{R}$.
% Indeed, it is a power series featuring nonnegative coefficients, which is dominated by $
%   \sum_{n \geq 1} \tfrac{x^n}{n!} \log{(n+\tfrac{m}2)}
% $. By the ratio test, the dominating
% series has infinite radius:
% \lim_{n\to\infty}
%   \biggl\lvert
%     \frac{
%       n! x^{n+1} \log{(n + 1 + \tfrac{m}2)}
%     }{
%       x^n \log{(n + \tfrac{m}2)} (n+1)!
%     }
%   \biggr\rvert
%   = \lim_{n\to\infty}
%     \lvert x \rvert 
%     \biggl\lvert
%       \frac{
%         \log{(n + 1 + \tfrac{m}2)}
%       }{
%         \log{(n + \tfrac{m}2)} (n+1)
%       }
%     \biggr\rvert
%   % = \lim_{n\to\infty}
%   %   \lvert x \rvert 
%   %   \biggl\lvert
%   %     \frac{
%   %       n + \tfrac{m}2
%   %     }{
%   %       (n + 1 + \tfrac{m}2)((n+1) + (n + \tfrac{m}2) \log{(n + \tfrac{m}2)})
%   %     }
%   %   \biggr\rvert
%   = 0 < 1
%   \,, $$
% since
% \frac{\log{(x + a + 1)}}{x \log{(x + a)}}
%   \sim \frac{\log{(x+1)}}{(x-a) \log{x}}
%   \sim \frac{\log{(x+1)}}{x \log{x}}
%   \sim \frac{\tfrac1{(x+1)}}{1 + \log{x}}
%   \to 0
% A theorem from calculus states, that the formal series derivative (integral)
% coincides with the derivative (integral) of the function, corresponding to
% the power series (everywhere on the convergence region). And the convergence
% regions of derivative (integral) coincide with the region of the original
% power series.
Formally differentiating the convergent power series within $g_m$ yields
\begin{equation}  \label{eq:g-m-deriv}
  \frac{d}{d x} g_m(x)
    % = e^{-x} \sum_{n\geq 1} \frac{x^{n-1}}{(n-1)!} \psi(n + \tfrac{m}2) - g_m(x)
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \psi(n + \tfrac{m}2 + 1) - g_m(x)
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} (
    %     \psi(n + \tfrac{m}2 + 1) - \psi(n + \tfrac{m}2)
    % )
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \tfrac1{n + \tfrac{m}2}
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} (n + \tfrac{m}2)^{-1}
    = x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
        \frac{x^{n + \tfrac{m}2}}{n!} (n + \tfrac{m}2)^{-1}

@@

From $
    e^t  = \sum_{n\geq 0} \tfrac{t^n}{n!}
$ on $\mathbb{R}$ and $
  \tfrac{x^\alpha}{\alpha}
    = \int_0^x t^{\alpha-1} dt
$ for $\alpha\neq 0$ we get
% by the monotone convergence theorem on ([0, x], \mathcal{B}([0, x]), dx)
\begin{align}  \label{eq:g-m-deriv-int-2}
  \eqref{eq:g-m-deriv}
    % = x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
    %     \frac{x^{n + \tfrac{m}2}}{n!} (n + \tfrac{m}2)^{-1}
    &= x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
        \frac1{n!} \int_0^x t^{n + \tfrac{m}2 - 1} dt
    %     \int_0^x t^{\tfrac{m}2 - 1} \sum_{n\geq 0} \frac{t^n}{n!} dt
    &= x^{-\tfrac{m}2} e^{-x}
        \int_0^x t^{\tfrac{m}2 - 1} e^t dt

@@

\end{align}
Substitution $u^2 = t$ on $[0, \infty]$ with $2u du = dt$ yields
\begin{equation}  \label{eq:g-m-deriv-int-3}
  \frac{d}{d x} g_m(x)

@@

    = 2 x^{-\tfrac{m}2} e^{-x}

@@

@@

In particular, the derivative of \eqref{eq:g-m-series} for $m=1$ is
\begin{equation}  \label{eq:g-m-deriv-one}
  \frac{d}{d x} g_1(x)
    = 2 \frac{F(\sqrt{x})}{\sqrt{x}}
using Dawson's integral $
  F
  \colon x \mapsto e^{-x^2} \int_0^x e^{u^2} du
Hence, the derivative of the expectation in \eqref{eq:improper-kl-div-real} with respect
to $\alpha$ is
$$
\frac{d}{d \alpha} g_1(\tfrac1{2\alpha})
  = -\frac1{2 \alpha^2} g_1'(\tfrac1{2\alpha})
  % = -\frac1{2 \alpha^2} 2 \tfrac{F(\sqrt{\tfrac1{2\alpha}})}{\sqrt{\tfrac1{2\alpha}}}
  % = -\frac1{2 \alpha^2} 2 \tfrac{F(\tfrac1{\sqrt{2\alpha}})}{\tfrac1{\sqrt{2\alpha}}}
  = -\frac1{\alpha} \sqrt{\frac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
  \,. $$
Finally, since $\alpha$ is nonnegative, it is typically parameterized via its logarithm.
The gradient of \eqref{eq:improper-kl-div-real} w.r.t $\log \alpha$ is

@@

  \frac{d \eqref{eq:improper-kl-div-real}}{d\log \alpha}
  % \frac{d}{d\log \alpha}
  %   \frac12 \mathbb{E}_{z \sim \mathcal{N}(0,1)}
  %     \log \lvert z + \tfrac1{\sqrt{\alpha}} \rvert^2
    % = \frac12 \frac{d\alpha}{d\log \alpha}
    %   \frac{d}{d\alpha} \bigl(\mathbb{E}\cdots \bigr)
    % = - \frac\alpha2 \tfrac1{\alpha}
    %   \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})

>>>>

% applications of cvnn
Complex-valued nature of the data in acoustic and radio signal processing has been
the main driver behind the adoption of $\cplx$-valued neural networks ($\cplx$VNN).
\citet{hirose_complex-valued_2009} argues that the combined phase-magnitude effect of
$\cplx$-valued transformations removes the excess degrees of freedom, that cause
degenerate transformations in $\real$-valued networks with twice the feature dimensions.
Their study demonstrates superiority of $\cplx$VNN in landmine detection using ground
penetrating radar imaging. Other examples, where $\cplx$-valued networks have outperformed
$\real$-valued networks, include magnetic resonance \citep{hui_mri_1995,wang_deepcomplexmri_2020}
and radar imaging \citep{haensch_complex-valued_2010,zhang_complex-valued_2017}, music
transcription and spectral speech modelling \citep{wisdom_full-capacity_2016,trabelsi_deep_2018},
and wireless signal classification \citep{yang_complex_2020}. \citet{tarver_design_2019}
have lowered the out-of-band power leakage with a $\cplx$-valued network for digital
signal predistortion. The networks have also been applied to non-$\cplx$-valued domains,
such as image classification \citep{popa_complex-valued_2017}, sequence modelling
\citep{danihelka_associative_2016}, and motion prediction \citep{wolter_complex_2018},
and for stabilizing back-propagation in RNN \citep{wisdom_full-capacity_2016}.

@@

\citet{kharitonov_variational_2018} addressed theoretical issues with improper prior
$\pi$ in Sparse VD, emphasized by \citet{hron_variational_2018}, and proposed
\emph{Automatic Relevance Determination} Variational Dropout, by replacing $\pi(\omega_j)$
with a proper Gaussian prior $
  \pi_\lambda(\omega_j) = \mathcal{N}(\omega_j \vert\, 0, \tau_j^{-1})
$ with learnable precision $\tau_j > 0$ \citep{neal_bayesian_1996}. This recast the VD
as the \emph{Empirical Bayes} approach, which performs Bayesian Inference over $\omega$,
but uses Maximum Likelihood estimates for the hyper-parameters $\lambda$,
\citep{mackay_bayesian_1994}. Maximizing \eqref{eq:sgvb_estimator} over $\tau$, holding
other parameters fixed, yields $
  \tau^*_j = {(\mu_j^2 + \sigma^2_j)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  K(\alpha)
    = \frac12 \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}
% \begin{align}
%   \pi(\omega)
%     &
%     = \prod_J \pi(s_J)
%       \prod_{j\in J} \pi(\omega_j \vert s_J)
%     \propto \prod_J \frac1{\lvert s_J \rvert}
%       \prod_{j\in J} \mathcal{N}(\omega_j \vert 0, s_J^2)
%     \,, \\
%   q_\theta(\omega)
%     &
%     % = \prod_J q_\theta(s_J)
%     %   \prod_{j\in J} q_\theta(\omega_j \vert s_J)
%     = \prod_J \mathcal{N}(s_J \vert \mu_J, \alpha_J \mu_J^2)
%       \prod_{j\in J} \mathcal{N}(\omega_j \vert \mu_j s_J, \sigma_j^2 s_J^2)
%     \,.
% \end{align}

@@

% Although $\cplx$-valued dense layers and back-propagation have been around since early
% 1990-s, one of the earliest accounts
Development of deep $\cplx$-valued networks has been active. \citet{haensch_complex-valued_2010}
put forward $\cplx$-valued convolutional networks, \citet{guberman_complex_2016} and
\citet{popa_complex-valued_2017} developed modifications of pooling, \citet{arjovsky_unitary_2016}
and \citet{wisdom_full-capacity_2016} proposed $\cplx$-valued RNNs with unitary recurrent
transition matrices, and \citet{danihelka_associative_2016} developed $\cplx$-valued
holographic representations for LSTMs. More recently \citet{trabelsi_deep_2018} proposed
$\cplx$-valued batch-normalization and weight initialization, \citet{wolter_complex_2018}
investigated different $\cplx$-valued gating mechanisms for RNNs, and \citet{yang_complex_2020}
proposed $\cplx$-valued self-attention and complex transformer architecture. It merits
noting that \citet{gaudet_deep_2018} generalized $\cplx$VNN further to deep quaternion-valued
networks, and \citet{vecchi_compressing_2020} studied sparsity inducing regularizers for
them.
% complex-valued distributions \citep{pav_moments_2015,taubock_complex-valued_2012},
% and \citep{karseras_caution:_2014}

@@

% Prior experiments demonstrated that the $\cplx$-valued variational sparsification
% what can be expected from prior research.
MusicNet is a corpus of $330$ annotated classical music recordings used for learning
feature representations for music transcription tasks \citep{thickstun_learning_2017}.
\citet{trabelsi_deep_2018} have proposed a $1d$ VGG-like $\cplx$-valued network that
surpassed a similar $\real$-valued network and achieved $72.9\%$ pooled \emph{Average
Precision} on this dataset. Recently \citet{yang_complex_2020} have reported $74.2\%$
AP with a $\cplx$-valued transformer, \citet{thickstun_invariances_2018} have achieved
$77.3\%$ with a four-layer $\real$-valued network on $\log$-spaced spectrogram, and
\citet{draguns_residual_2020} report $78.0\%$ with a residual-shuffle-exchange network.

@@

In this appendix we study the approximation proposed by \citet{molchanov_variational_2017}
for the KL divergence term \eqref{eq:improper-kl-div-real}. We show empirically that
the gradient feedback provided by the approximation is sufficiently close to $
  \tfrac{d}{d\log \alpha} K(\alpha)
$, making it a viable alternative to exact gradients, computed in \citet{hron_variational_2018}.
% The derivation in this section different from \citet{lapidoth_capacity_2003}.
% The results are identical: they use DCT to establish interchangeability of limits
% (sum-integral-derivative), whilst we allude to calculus theorems on differentiation
% and integration of power series within the radius of their convergence (which are
% essentially corollaries to the DCT).

% \citet{pav_moments_2015} correctly notice that \cite[p. 2466]{lapidoth_capacity_2003} has a missing $\log{2}$ term.

@@

$, the random variable $W = \sum_i (\mu_i + z_i)^2$ has non-central $\chi^2$ distribution
with shape $m$ and non-centrality parameter $\lambda = \sum_i \mu_i^2$, i.e. $
  W\sim \chi^2_m(\lambda)
$. The divergence \eqref{eq:improper-kl-div-real} has the form
\begin{equation}  \label{eq:kl-div-appendix}
  K(\alpha)
    \propto \frac12 \mathbb{E}_{
        W\sim \chi^2_1\bigl(\tfrac1{\alpha}\bigr)
      } \log W
    \tag{\ref{eq:improper-kl-div-real}'}
$W$ can alternatively be represented as a Poisson mixture of ordinary $\chi^2$
distributions: if $Z_{\mid J} \sim \chi^2_{k + 2J}$ for $
  J \sim \mathcal{P}ois(\tfrac{\lambda}2)
$ then $W \sim Z$. Therefore, expanding the conditional expectation gives
\begin{equation}  \label{eq:cond_exp_expansion}
  \mathbb{E}_{W\sim \chi^2_k(\lambda)} \log W
    = \mathbb{E}\Bigl(
        \mathbb{E} \bigl( \log W \mid J)
    = \mathbb{E}_{J\sim {\mathcal{P}ois}(\tfrac\lambda2)}
        \mathbb{E}_{W\sim \chi^2_{m + 2 J}} \log W

@@

Since $\chi^2_\nu$ is Gamma distribution $\Gamma(\tfrac\nu2, \tfrac12)$, its logarithmic
moment $
  \mathbb{E}_{W\sim \chi^2_\nu} \log W
$ is $
  \psi\bigl(\tfrac\nu{2}\bigr) - \log \tfrac12
$, where $\psi(x)$ is the digamma function. Hence
\begin{equation}  \label{eq:exp_digamma}
  \mathbb{E}_{W\sim \chi^2_k(\lambda)} \log W
    = \log2 + \mathbb{E}_{J\sim {\mathcal{P}ois}(\tfrac\lambda2)}
      \psi\bigl(\tfrac{m + 2 J}{2}\bigr)
    % = \log2
    % + \sum_{j\geq 0}
    %     e^{- \tfrac{\lambda}2}
    %     \frac{\bigl(\tfrac{\lambda}2\bigr)^j}{j!}
    %   \psi\bigl(\tfrac{m+2J}{2}\bigr)
We can use REINFORCE method to compute the derivative of \eqref{eq:exp_digamma} with respect
to $\lambda$ \citep{williams_simple_1992}. Since Poisson $\log$-probability is $
  \log p_\lambda(J)
    = - \tfrac{\lambda}2
      + J \log \tfrac{\lambda}2
      - \log J!
$, we get
\begin{equation}  \label{eq:exp_reinforce}
  \frac{d \eqref{eq:exp_digamma}}{d \lambda}
    =
    \mathbb{E}_{J\sim {\mathcal{P}ois}(\tfrac\lambda2)}
      \psi\bigl(\tfrac{m + 2 J}{2}\bigr)
      \tfrac{d}{d \lambda} \log p_\lambda(J)
    % = \mathbb{E}_{J}
    %   \bigl(
    %     \tfrac2\lambda J - 1
    %   \bigr)
    %   \frac{\psi\bigl(\tfrac{m + 2 J}{2}\bigr)}2
    = \frac12 g_m\bigl(\tfrac\lambda2\bigr)

@@

where
\begin{equation}  \label{eq:g_m_series}

@@

    =
    \mathbb{E}_{J\sim {\mathcal{P}ois}(x)}
      \bigl( \tfrac{J}x - 1 \bigr)
      \psi\bigl(\tfrac{m + 2 J}{2}\bigr)
    =
    e^{- x} \sum_{j \geq 0}
      \frac{x^j}{j!} \bigl( \tfrac{J}x - 1 \bigr)
      \psi\bigl(\tfrac{m + 2 j}{2}\bigr)

@@

By manipulating the partial sums within $g_m(x)$ and making use of the property $
  \psi(x + 1) = \psi(x) + \tfrac1x
  % \psi(1 + J + \tfrac{m}2) = \tfrac2{m + 2J} + \psi(J + \tfrac{m}2)
$ of the digamma funciton for $x > 0$, we get
\begin{equation}  \label{eq:g_m_series_int}
    = e^{- x} \sum_{j \geq 0}
      \frac{x^j}{j!} \frac1{j + \tfrac{m}2}
    = e^{- x} x^{- \tfrac{m}2} \sum_{j \geq 0}
      \frac1{j!}
      \frac{x^{j + \tfrac{m}2}}{j + \tfrac{m}2}
    = e^{- x} x^{- \tfrac{m}2} \sum_{j \geq 0}
      \frac1{j!} \int_0^x t^{j + \tfrac{m}2 - 1} dt

@@

Since for any $J$ the functions $
  t \mapsto \sum_{j=0}^J \frac1{j!} t^{j + \tfrac{m}2 - 1}
$ are positive on $(0, x)$ and upper bounded by integrable $
  t^{\tfrac{m}2 - 1} e^t
$, which turns out to be the limit of the partial sums, the Dominated Convergence
theorem implies
\begin{equation}  \label{eq:g_m_integral}
    = e^{- x} x^{- \tfrac{m}2}
      \int_0^x \sum_{j \geq 0}
        \frac1{j!} t^{j + \tfrac{m}2 - 1} dt
    = e^{- x} x^{- \tfrac{m}2}
      \int_0^x e^t t^{(\tfrac{m}2 - 1)} dt
Substituting $u^2 = t$ on $[0, \infty]$ with $2u du = dt$ and letting $
  F_m
  \colon x \mapsto e^{-x^2} \int_0^x u^{m - 1} e^{u^2} du
$ yields
\begin{equation}  \label{eq:g_m_integral-sub}
  \frac{d \eqref{eq:exp_digamma}}{d \lambda}
    = \frac12 g_m\bigl(\tfrac\lambda2\bigr)

@@

    = e^{-x}
      x^{-\tfrac{m}2}

@@

        \bigg\vert_{x=\tfrac\lambda2}
    % = e^{-x}
    %   \frac1{\sqrt{x}}
    %   \int_0^{\sqrt{x}} e^{u^2} du
    %     \bigg\vert_{x=\tfrac\lambda2}
    = \Bigl(\sqrt{\tfrac2\lambda}\Bigr)^m F\Bigl(\sqrt{\tfrac\lambda2}\Bigr)

@@

Since $\alpha$ is non-negative, it is typically parameterized via its logarithm, whence
the derivative of \eqref{eq:kl-div-appendix} with respect to $\log\alpha$ follows from
\eqref{eq:g_m_integral-sub} for $m=1$ and $\lambda = \tfrac1\alpha$:

@@

  \frac{d K(\alpha)}{d\log \alpha}
    % = \tfrac12 \sqrt{\tfrac2\lambda} F\Bigl(\sqrt{\tfrac\lambda2}\Bigr)
    % for \lambda = \tfrac1\alpha
    %   \tfrac{d\lambda}{d\alpha} = - \alpha^{-2}
    %   \tfrac{d\alpha}{d\log \alpha} = \alpha
    % = - \alpha \frac1{\alpha\sqrt{2\alpha}} F\Bigl(\tfrac1{\sqrt{2\alpha}}\Bigr)
